## Functions and Conditionals

### Conditionals

The basic syntax for an if statement goes like this:

```{r}
if (logical statement) {
  do something
} else {
  do something else
}
```

The logical statements should boil down to a boolean; for example:

```{r}

```

We can also add else if statements in the middle if we need more options.

```{r}

```

Neat! We can also nest conditionals just like we do in Python.

```{r}

    
```

Neat! Now, let's look at another core part of R programming - Loops!

### Loops

```{r}

```

We can also do nifty stuff with our vectors, such as apply changes to specific indices of our data.

```{r}

```

We can loop over more than just lists:

```{r}

```

Like Python, we have while loops in R:

```{r}

```

## Functions

When should we use functions?

```{r}
our_function <- function(optional_params){
  # do some stuff
}
```

An example:

```{r}

```

This is all cool.. but it didn't do anything. We need to call the function!

```{r}

```

Notice that in the background here, we are looping over each item in the array, but we never write a for loop! Keep this in mind, since when you pass in a large data set, the changes made by functions aren't quite immediate - the computer does have to go over each item in your data set, and change them one at a time.

A shortcut of sorts is to use the `apply()` function, which operates very similarly to Python's `map()`. Note that `apply` syntax goes as follows: `apply(X, MARGIN, FUN)` where X is the data set to use, MARGIN is specification for rows or columns, and FUN is the function to apply.

```{r}

```

It is common to use `mean`, `median`, `sum`, `min`, or `max` with `apply`, but we can also use our user-defined functions as well. Note that apply() only works with matrices.

```{r}

```

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:

    ```{r}
    
    
    # First Code
    rps <- function(p1, p2){
  valid_moves <- c("rock", "paper", "scissors")
  if (!all(p1 %in% valid_moves) | !all(p2 %in% valid_moves)) {
    stop("Invalid input: Moves must be 'rock', 'paper', or 'scissors'.")
  }
  
  winning_combos <- list(rock = "scissors", paper = "rock", scissors = "paper")
  if (winning_combos[[p1]] == p2) {
    "Player 1 won!"
  } else if (p1 == p2) {
    "Draw!"
  } else {
    "Player 2 won!"
  }
}

  # Second Code
  good_bad_ideas <- function(x) {
  num_good_ideas <- sum(x == "good")
  
  if (num_good_ideas == 0) {
    "Fail!"
  } else if (num_good_ideas <= 2) {
    "Publish!"
  } else {
    "I smell a series!"
  }
}

  


  # Third Code 
pofi <- function(n) {
  if (!is.numeric(n) | n < 0) {
    stop("Invalid input: Exponent must be a non-negative integer.")
  }
  
  pattern <- c(1, "i", -1, "-i")
  
  return(pattern[(n %% 4) + 1])
}



# fourth code 
sum_array <- function(a) {
  total_sum <- 0
  
  for (number in a) {
    total_sum <- total_sum + number
  }
  return(total_sum)
}



# fifth code 
even_or_odd <- function(number) {
  if (number %% 2 == 0) {
    "Even"
  } else {
    "Odd"
  }
}



    ```

2.  Create a function that checks an input of type `int` and returns fizz if the number is a multiple of 3, buzz if the number is a multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

    ```{r}
fizzbuzz <- function(number) {

  is_multiple_of_3 <- number %% 3 == 0
  is_multiple_of_5 <- number %% 5 == 0

  if (is_multiple_of_3 & is_multiple_of_5) {
    "fizzbuzz"
  } else if (is_multiple_of_3) {
    "fizz"
  } else if (is_multiple_of_5) {
    "buzz"
  } else {
    "neither a fizz not a buzz"
  }
}

# Example usage
numbers <- c(3, 5, 15, 4, 7)
fizzbuzz_results <- sapply(numbers, fizzbuzz)
print(fizzbuzz_results)
    ```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends each level by ascending a flight of stairs and leaping off the top to arrive at the castle at the end. (See [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can imagine that each of the bricks of the stairs is a character X. Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4, we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX

    ```{r}
    # You may feel more comfortable figuring this out in Python first, then translating into R later.
    
    
marioStairs <- function(num_stairs) {

  if (!is.numeric(num_stairs) || num_stairs <= 0) {
    stop("Invalid input: Number of stairs must be a positive integer.")
  }
  
  for (i in 1:num_stairs) {
    for (j in 1:i) {
      cat("X", sep = "")
    }
    cat("\n")
  }
}

# Example usage
num_stairs <- 6
marioStairs(num_stairs)
    ```
